---
title: "WNBA and NBA Final"
subtitle: "By: Louise and Maia"
runtime: shiny
output:
  flexdashboard::flex_dashboard:
    theme: 
      bg: "#fcfbf8"
      fg: "#171c19" 
      primary: "#e9b0fc"
      navbar-bg: "#e9b0fc"
---

```{r global, include= FALSE, message=FALSE}
library(readr)
library(dplyr)
library(ggplot2)
library(shiny)
library(tidyverse)
library(shinydashboard)
library(shinyWidgets)
library(rvest)
library(httr)
library(purrr)
library(leaflet)

```

About
==========================

Column {data-width = 400}
-------------------------

For our final project, we aimed to scrape data from the WNBA to uncover insights into the statistics behind women's basketball. Using the skills and tools we have acquired in Data Science 2, we collected data on various aspects of the WNBA in 2023. Our analysis includes a comparison of games played, player positions, teams, and the relationship between player height and age.

Additionally, we incorporated a comparative analysis with the NBA from 2018 to 2024. This comparison examines metrics such as the number of active players, average age, total capital, team spending limits, and current expenditure on players. We utilized multiple data sources to ensure a comprehensive analysis, scraping data from various websites, player profiles, and financial reports.

Through this project, we aim to highlight key statistical differences and similarities between the WNBA and NBA, providing a thorough overview of the landscape of professional basketball for both women and men. We have included visualizations such as bar charts and scatter plots to clearly present our findings. Furthermore, we created interactive dashboards that allow users to explore the data in more detail, offering insights into specific teams and players.

We hope you find our exploration of these sports statistics both informative and engaging, shedding light on the dynamics of professional basketball across genders.

```{r}
library(leaflet)
library(dplyr)

# Create a data frame with the latitude and longitude for each WNBA stadium
wnba_stadiums <- data.frame(
  team = c("Atlanta Dream", "Chicago Sky", "Connecticut Sun", "Dallas Wings", "Indiana Fever", 
           "Las Vegas Aces", "Los Angeles Sparks", "Minnesota Lynx", "New York Liberty", 
           "Phoenix Mercury", "Seattle Storm", "Washington Mystics"),
  lat = c(33.7573, 41.8807, 41.4896, 32.7905, 39.7640, 36.0905, 34.0430, 44.9795, 
          40.7505, 33.4456, 47.6222, 38.8684),
  lon = c(-84.3963, -87.6742, -72.0863, -96.8103, -86.1555, -115.1750, -118.2673, 
          -93.2760, -73.9934, -112.0712, -122.3545, -76.9930)
)

# Group teams that share the same stadium together
wnba_stadiums <- wnba_stadiums %>%
  group_by(lat, lon) %>%
  summarise(team = paste(team, collapse = "<br>"))

# Create a pink basketball icon
basketball_icon <- makeIcon(
  iconUrl = "https://www.prosportstickers.com/wp-content/uploads/2022/10/woman-basketball-ball-emoji.png", # URL to the pink basketball emoji
  iconWidth = 28,
  iconHeight = 28
)

# Create a leaflet map
leaflet(wnba_stadiums) %>%
  addTiles() %>%
  addMarkers(~lon, ~lat, icon = basketball_icon, popup = ~team)
```


#### Background Information

The WNBA regular season consists of 40 games, which concludes in September, compared to the NBA's 82-game regular season. The NBA season starts on Tuesday, October 24, while the WNBA season begins in May, coinciding with the NBA playoffs. The WNBA also uses a slightly smaller ball, measuring 28.5 inches in circumference, compared to the NBA's 29.5-inch ball. Additionally, the style of play differs significantly between the two leagues, with the WNBA often emphasizing teamwork, fundamentals, and strategic play, while the NBA is known for its fast pace, athleticism, and individual star power.

#### Data Source
We drew data from these three websites:
https://www.espn.com, https://hoopshype.com, and http://spotrac.com

Demographics of the WNBA
==========================

Inputs {.sidebar}
-------------------------

```{r, include=FALSE, message=FALSE}

url <- "https://www.espn.com/wnba/team/roster/_/name/chi/chicago-sky"
robotstxt::paths_allowed(url)  

namechi <- read_html(url) |>
  html_nodes(".inline .AnchorLink") |>
  html_text()
agechi <- read_html(url) |>
  html_nodes(".Table__TD:nth-child(3) .inline") |>
  html_text()
heightchi <- read_html(url) |>
  html_nodes(".Table__TD:nth-child(4) .inline") |>
  html_text()


urlatl <- "https://www.espn.com/wnba/team/roster/_/name/atl/atlanta-dream"
robotstxt::paths_allowed(urlatl)  
nameatl <- read_html(urlatl) |>
  html_nodes(".inline .AnchorLink") |>
  html_text()
ageatl <- read_html(urlatl) |>
  html_nodes(".Table__TD:nth-child(3) .inline") |>
  html_text()
heightatl <- read_html(urlatl) |>
  html_nodes(".Table__TD:nth-child(4) .inline") |>
  html_text()


urlconn <- "https://www.espn.com/wnba/team/roster/_/name/conn/connecticut-sun"
robotstxt::paths_allowed(urlconn)  
nameconn <- read_html(urlconn) |>
  html_nodes(".inline .AnchorLink") |>
  html_text()
ageconn <- read_html(urlconn) |>
  html_nodes(".Table__TD:nth-child(3) .inline") |>
  html_text()
heightconn <- read_html(urlconn) |>
  html_nodes(".Table__TD:nth-child(4) .inline") |>
  html_text()


urldal <- "https://www.espn.com/wnba/team/roster/_/name/dal/dallas-wings"
robotstxt::paths_allowed(urldal)  
namedal <- read_html(urldal) |>
  html_nodes(".inline .AnchorLink") |>
  html_text()
agedal <- read_html(urldal) |>
  html_nodes(".Table__TD:nth-child(3) .inline") |>
  html_text()
heightdal <- read_html(urldal) |>
  html_nodes(".Table__TD:nth-child(4) .inline") |>
  html_text()

urlind <- "https://www.espn.com/wnba/team/roster/_/name/ind/indiana-fever"
robotstxt::paths_allowed(urlind)  
nameind <- read_html(urlind) |>
  html_nodes(".inline .AnchorLink") |>
  html_text()
ageind <- read_html(urlind) |>
  html_nodes(".Table__TD:nth-child(3) .inline") |>
  html_text()
heightind <- read_html(urlind) |>
  html_nodes(".Table__TD:nth-child(4) .inline") |>
  html_text()

urllv <- "https://www.espn.com/wnba/team/roster/_/name/lv/las-vegas-aces"
robotstxt::paths_allowed(urllv)  
namelv <- read_html(urllv) |>
  html_nodes(".inline .AnchorLink") |>
  html_text()
agelv <- read_html(urllv) |>
  html_nodes(".Table__TD:nth-child(3) .inline") |>
  html_text()
heightlv <- read_html(urllv) |>
  html_nodes(".Table__TD:nth-child(4) .inline") |>
  html_text()

urlla <- "https://www.espn.com/wnba/team/roster/_/name/la/los-angeles-sparks"
robotstxt::paths_allowed(urlla)  
namela <- read_html(urlla) |>
  html_nodes(".inline .AnchorLink") |>
  html_text()
agela <- read_html(urlla) |>
  html_nodes(".Table__TD:nth-child(3) .inline") |>
  html_text()
heightla <- read_html(urlla) |>
  html_nodes(".Table__TD:nth-child(4) .inline") |>
  html_text()

urlmin <- "https://www.espn.com/wnba/team/roster/_/name/min/minnesota-lynx"
robotstxt::paths_allowed(urlmin)  
namemin <- read_html(urlmin) |>
  html_nodes(".inline .AnchorLink") |>
  html_text()
agemin <- read_html(urlmin) |>
  html_nodes(".Table__TD:nth-child(3) .inline") |>
  html_text()
heightmin <- read_html(urlmin) |>
  html_nodes(".Table__TD:nth-child(4) .inline") |>
  html_text()

urlny <- "https://www.espn.com/wnba/team/roster/_/name/ny/new-york-liberty"
robotstxt::paths_allowed(urlny)  
nameny <- read_html(urlny) |>
  html_nodes(".inline .AnchorLink") |>
  html_text()
ageny <- read_html(urlny) |>
  html_nodes(".Table__TD:nth-child(3) .inline") |>
  html_text()
heightny <- read_html(urlny) |>
  html_nodes(".Table__TD:nth-child(4) .inline") |>
  html_text()

urlphx <- "https://www.espn.com/wnba/team/roster/_/name/phx/phoenix-mercury"
robotstxt::paths_allowed(urlphx)  
namephx <- read_html(urlphx) |>
  html_nodes(".inline .AnchorLink") |>
  html_text()
agephx <- read_html(urlphx) |>
  html_nodes(".Table__TD:nth-child(3) .inline") |>
  html_text()
heightphx <- read_html(urlphx) |>
  html_nodes(".Table__TD:nth-child(4) .inline") |>
  html_text()

urlsea <- "https://www.espn.com/wnba/team/roster/_/name/sea/seattle-storm"
robotstxt::paths_allowed(urlsea)  
namesea <- read_html(urlsea) |>
  html_nodes(".inline .AnchorLink") |>
  html_text()
agesea <- read_html(urlsea) |>
  html_nodes(".Table__TD:nth-child(3) .inline") |>
  html_text()
heightsea <- read_html(urlsea) |>
  html_nodes(".Table__TD:nth-child(4) .inline") |>
  html_text()

urlwsh <- "https://www.espn.com/wnba/team/roster/_/name/wsh/washington-mystics"
robotstxt::paths_allowed(urlwsh)  
namewsh <- read_html(urlwsh) |>
  html_nodes(".inline .AnchorLink") |>
  html_text()
agewsh <- read_html(urlwsh) |>
  html_nodes(".Table__TD:nth-child(3) .inline") |>
  html_text()
heightwsh <- read_html(urlwsh) |>
  html_nodes(".Table__TD:nth-child(4) .inline") |>
  html_text()
nameconn <- append(nameconn, "Tiffany Hayes")
ageconn <- append(ageconn, 33)
heightconn <- append(heightconn, "5' 10\"")
playerfull <- c(nameatl,namechi,nameconn,namedal,nameind,namelv,namela,namemin,nameny,namephx, namesea,namewsh)
agefull <- c(ageatl,agechi,ageconn,agedal,ageind,agelv,agela,agemin,ageny,agephx, agesea,agewsh)
heightfull <-c(heightatl,heightchi,heightconn,heightdal,heightind,heightlv,heightla,heightmin,heightny,heightphx, heightsea,heightwsh) 
```

```{r, include=FALSE, message=FALSE}
scrape_teamdata <- function(year){
  urlteam <-str_c("https://www.espn.com/wnba/standings/_/season/",year, sep = "")

robotstxt::paths_allowed(urlteam)  

teamm <- read_html(urlteam) |>
  html_nodes(".hide-mobile .AnchorLink") |>
  html_text()

wins <- read_html(urlteam) |>
  html_nodes(".Table__TD:nth-child(1) .stat-cell") |>
  html_text()
wins

losses <- read_html(urlteam) |>
  html_nodes(".Table__TD:nth-child(2) .stat-cell") |>
  html_text()

percent_won <- read_html(urlteam) |>
  html_nodes(".Table__TD:nth-child(3) .stat-cell") |>
  html_text()

home_record <- read_html(urlteam) |>
  html_nodes(".Table__TD:nth-child(5) .stat-cell") |>
  html_text()

away_record <- read_html(urlteam) |>
  html_nodes(".Table__TD:nth-child(6) .stat-cell") |>
  html_text()

conf_record <- read_html(urlteam) |>
  html_nodes(".Table__TD:nth-child(7) .stat-cell") |>
  html_text()

points_per_game <- read_html(urlteam) |>
  html_nodes(".Table__TD:nth-child(8) .stat-cell") |>
  html_text()

tibble(teamm, wins, losses, percent_won, home_record, away_record, conf_record, points_per_game) |>
  mutate(wins=as.double(wins), losses= as.double(losses), points_per_game= as.double(points_per_game), percent_won= as.double(percent_won)) |>
  mutate(year = year, team= teamm) |>
  select(-teamm)
}


scrape_teamdata(2023)


teamstatdata <- bind_rows(scrape_teamdata(2016),
          scrape_teamdata(2017),
          scrape_teamdata(2018), 
          scrape_teamdata(2019), 
          scrape_teamdata(2020), 
          scrape_teamdata(2021), 
          scrape_teamdata(2022), 
          scrape_teamdata(2023))

teamstatdata |>
  count(year) 


scrape_playerdata <- function(year){
  urlplayer <-str_c("https://www.espn.com/wnba/stats/player/_/season/",year,"/seasontype/2", sep = "")

robotstxt::paths_allowed(urlplayer)  

playeryear <- read_html(urlplayer) |>
  html_nodes(".mr7 .AnchorLink") |>
  html_text() 

pointsyear <- read_html(urlplayer) |>
  html_nodes(".Table__TD:nth-child(4)") |>
  html_text()

games_playedyear <- read_html(urlplayer) |>
  html_nodes(".Table__Scroller .Table__TD:nth-child(2)") |>
  html_text() 

rankyear <- read_html(urlplayer) |>
  html_nodes(".Table--fixed-left .Table__TH:nth-child(1) div , .Table--fixed-left .Table__TD:nth-child(1)") |>
  html_text() |>
  tail(n=50)

positionyear <- read_html(urlplayer) |>
  html_nodes(".position") |>
  html_text()|>
  head(n=50)

teamyear <- read_html(urlplayer) |>
  html_nodes(".athleteCell__teamAbbrev") |>
  html_text()|>
  head(n=50)

primary_data <- tibble(playeryear, pointsyear, games_playedyear, rankyear, positionyear, teamyear)

dataage <- tibble(playerfull,agefull,heightfull) |>
  mutate(agefull=(as.double(agefull)-(2024-year)))

left_join(primary_data, dataage, 
                           by = join_by(playeryear== playerfull)) |>
  mutate(heightfull= str_replace(heightfull, '"', "")) |>
  mutate(heightfull= str_replace(heightfull, "'", "")) |>
  separate(heightfull, c("feet", "inches"), " ") |>
  mutate(feet= as.double(feet), inches= as.double(inches)) |>
  mutate(feet_in_inches= feet*12, inches_in_feet= inches/12) |>
  mutate(height_in = feet_in_inches+inches, 
         height_ft= feet+inches_in_feet) |>
  select(-feet, -inches, -feet_in_inches, -inches_in_feet) |>
  mutate(year=year)
}


scrape_playerdata(2023)


playerstatdata <- bind_rows(
          scrape_playerdata(2018), 
          scrape_playerdata(2019), 
          scrape_playerdata(2020), 
          scrape_playerdata(2021), 
          scrape_playerdata(2022), 
          scrape_playerdata(2023))

playerstatdata |>
  count(year) 
```

```{r, echo= FALSE}
inputPanel(
      selectInput(inputId="year", label= "Age Response Variable",
                  choices = c("Games Played" = "games_playedyear", 
                              "Position" = "positionyear", 
                              "Team" = "teamyear"),
                  selected = "teamyear"),
      selectInput("x", label = "Year",
                  choices = c("2018" = "2018", 
                              "2019" = "2019", 
                              "2020" = "2020", 
                              "2021"= "2021", 
                              "2022"= "2022", 
                              "2023"= "2023"),
                  selected = "2023")
)


inputPanel(
      selectInput(inputId = "y2", label = "Height Response Variable",
                  choices = c("Games Played" = "games_playedyear", 
                              "Position" = "positionyear", 
                              "Team" = "teamyear"),
                  selected = "teamyear"),
      selectInput("x2", label = "Year",
                  choices = c("2018" = "2018", 
                              "2019" = "2019", 
                              "2020" = "2020", 
                              "2021"= "2021", 
                              "2022"= "2022", 
                              "2023"= "2023"),
                  selected = "2023")
)

strong("In this section we hope to examine the demographics of the top 50 ranked players (ranked by points per game) for each year to answer the following questions: What ages play the most? What is the average height on each team? How does this affects their stats?")
       
strong("Key: C= Center
  F= Forward
  G= Gaurd")

p("One relationship that was interesting was that players who are in the forward position tend to be the youngest. There were no other clear relationships between age/height and stats.")


```


Column
-------------------------

```{r, echo= FALSE}
renderPlot({
 playerstatdata |> 
    drop_na() |>
    filter(year %in% c(input$x)) |>
    ggplot(aes(x=.data[[input$year]], y= agefull)) +
    geom_boxplot(color= "#fe6c5d") +
      coord_flip()+
    theme_classic() +
    labs(title= "How does age affect player stats?", y= "Age (years)", 
         x="Selected Stat")
  }) 


renderPlot({
 playerstatdata |> 
    drop_na() |>
    filter(year %in% c(input$x2)) |>
    ggplot(aes(x=.data[[input$y2]], y= height_in)) +
    geom_boxplot(color= "#fe6c5d") +
      coord_flip()+
    theme_classic() +
    labs(title= "How does height affect player stats?", y= "Height (in)", 
         x="Selected Stat")
  }) 

```



WNBA Team Statistics
==========================

Inputs {.sidebar}
-------------------------
```{r, echo= FALSE}
inputPanel(
      selectInput(inputId="xaxis", label= "Team Statistics Response Variable",
                  choices = c("Points per Game" = "points_per_game", 
                              "Percent Won" = "percent_won", 
                              "Number of Wins" = "wins", 
                              "Number of Losses"= "losses"),
                  selected = "points_per_game"),
      selectInput("yearteam", label = "Year",
                  choices = c("2018" = "2018", 
                              "2019" = "2019", 
                              "2020" = "2020", 
                              "2021"= "2021", 
                              "2022"= "2022", 
                              "2023"= "2023"),
                  selected = "2023")
)

inputPanel(
  selectInput(inputId="xaxis2", label= "Teams Over Time Response Variable",
                  choices = c("Points per Game" = "points_per_game", 
                              "Percent Won" = "percent_won", 
                              "Number of Wins" = "wins", 
                              "Number of Losses"= "losses"),
                  selected = "points_per_game"),
  selectInput(inputId="teamselect", label= "Team",
                  choices = c("Minnesota Lynx" = "Minnesota Lynx", 
                              "Los Angeles Sparks" = "Los Angeles Sparks", 
                              "New York Liberty" = "New York Liberty", 
                              "Chicago Sky"= "Chicago Sky", 
                              "Indiana Fever"= "Indiana Fever", 
                              "Atlanta Dream"= "Atlanta Dream", 
                              "Seattle Storm" = "Seattle Storm", 
                              "Phoenix Mercury" = "Phoenix Mercury", 
                              "Connecticut Sun" = "Connecticut Sun", 
                              "Washington Mystics"= "Washington Mystics", 
                              "Dallas Wings"= "Dallas Wings", 
                              "San Antonio Stars"= "San Antonio Stars", 
                              "Las Vegas Aces"= "Las Vegas Aces"),
                  selected = "Atlanta Dream")
)

strong("We also hoped to examine some team statistics as they evolve through the years. We hope to answer the questions: What teams did the best each year? By what variable? How have teams improved throughout the years?")

p("One interesting relationship we found in this section is that the number of losses and wins in 2020 seemed to dropped. This would make sense because there were less games, but interestingly this isn't the case for all teams, for example the Los Vegas Aces.")
```

Column
-------------------------
```{r, echo= FALSE}
renderPlot({
 teamstatdata |> 
    filter(year %in% c(input$yearteam)) |>
    mutate(team= fct_reorder(team,.data[[input$xaxis]])) |>
    ggplot(aes(x=team, y= .data[[input$xaxis]])) +
    geom_col(fill= "#fe6c5d") +
    coord_flip() +
    theme_classic()+
    labs(title= "Team Statistics", y= "Stat", x= "Team")
  }) 


```

```{r, echo= FALSE}
renderPlot({
 teamstatdata |> 
    filter(team %in% c(input$teamselect)) |>
    mutate(team= fct_reorder(team,.data[[input$xaxis2]])) |>
    ggplot(aes(x=year, y= .data[[input$xaxis2]])) +
    geom_point(color= "#fe6c5d")+
    geom_smooth(color= "pink", se= FALSE) +
    theme_classic() +
    labs(title= "Team Stats Througout the Years", x= "Year", y= "Response Variable")
  }) 

```

Column
-------------------------


How does the NBA differ from the WNBA?
==========================

```{r, include= FALSE, message=FALSE}
# NBA Data
scrape_nbadata <- function(start_year){
  url <-str_c("https://www.spotrac.com/nba/cap/_/year/", start_year, "/sort/cap_maximum_space", sep = "")

  nba <- read_html(url)

  tables <- html_nodes(nba, css = "table") 

  nba_data <- html_table(tables, header = TRUE, fill = TRUE)[[1]] 

  # Initialize year column
  nba_data$year <- NA

  # Initialize year and counter
  year <- start_year
  counter <- 0

  # Loop over rows
  for(i in seq_len(nrow(nba_data))) {
    # If the next rank is 1 and is not NA, increment the counter
    if(i < nrow(nba_data) && !is.na(nba_data$Rank[i+1]) && nba_data$Rank[i+1] == 1) {
      counter <- counter + 1
    }
    # Calculate the year based on the counter
    nba_data$year[i] <- start_year + counter
  }

  # Remove rows with "Totals" or "Averages"
  nba_data <- nba_data[!rowSums(sapply(nba_data, function(x) grepl("Totals|Averages", x))), ]

  nba_data <- nba_data %>%
    mutate(Team = str_sub(Team, start = 1, end = 3))
  return(nba_data)
}

scrape_nbadata(2023)

# Scrape data for all years
nba <- bind_rows(scrape_nbadata(2017),
          scrape_nbadata(2018),
          scrape_nbadata(2019),
          scrape_nbadata(2020),
          scrape_nbadata(2021),
          scrape_nbadata(2022),
          scrape_nbadata(2023))

# Rename columns
colnames(nba)[1] <- "rank"
colnames(nba)[2] <- "team"
colnames(nba)[4] <- "players_active"
colnames(nba)[5] <- "avg_age_team"
colnames(nba)[6] <- "total_cap_all"
colnames(nba)[7] <- "cap_space_all"
colnames(nba)[9] <- "active"

# Select columns and add new column
nba <- nba |>
  select (year, rank, team, players_active, avg_age_team, total_cap_all, cap_space_all, active) |>
  mutate(WM = "M")

nba
```

```{r, include=FALSE, message=FALSE}
# Create a data frame with team abbreviations, full names, and cities
nba_teams <- tibble(
  team = c("BOS", "BKN", "NYK", "PHI", "TOR", "CHI", "CLE", "DET", "IND", "MIL", "ATL", "CHA", "MIA", "ORL", "WAS", "DEN", "MIN", "OKC", "POR", "UTA", "GSW", "LAC", "LAL", "PHX", "SAC", "DAL", "HOU", "MEM", "NOP", "SAS"),
  team_name = c("Boston Celtics", "Brooklyn Nets", "New York Knicks", "Philadelphia 76ers", "Toronto Raptors", "Chicago Bulls", "Cleveland Cavaliers", "Detroit Pistons", "Indiana Pacers", "Milwaukee Bucks", "Atlanta Hawks", "Charlotte Hornets", "Miami Heat", "Orlando Magic", "Washington Wizards", "Denver Nuggets", "Minnesota Timberwolves", "Oklahoma City Thunder", "Portland Trail Blazers", "Utah Jazz", "Golden State Warriors", "LA Clippers", "Los Angeles Lakers", "Phoenix Suns", "Sacramento Kings", "Dallas Mavericks", "Houston Rockets", "Memphis Grizzlies", "New Orleans Pelicans", "San Antonio Spurs"),
  city = c("Boston", "Brooklyn", "New York", "Philadelphia", "Toronto", "Chicago", "Cleveland", "Detroit", "Indianapolis", "Milwaukee", "Atlanta", "Charlotte", "Miami", "Orlando", "Washington", "Denver", "Minneapolis", "Oklahoma City", "Portland", "Salt Lake City", "San Francisco", "Los Angeles", "Los Angeles", "Phoenix", "Sacramento", "Dallas", "Houston", "Memphis", "New Orleans", "San Antonio")
)

# Join NBA data with team names and cities
nba <- nba %>%
  left_join(nba_teams, by = "team")
```

```{r, include=FALSE, message=FALSE}
# WNBA Data
scrape_wnbadata <- function(start_year){
  url <-str_c("https://www.spotrac.com/wnba/cap/_/year/", start_year, "/sort/cap_maximum_space", sep = "")

  wnba <- read_html(url)

  tables <- html_nodes(wnba, css = "table") 

  wnba_data <- html_table(tables, header = TRUE, fill = TRUE)[[1]] 

  # Initialize year column
  wnba_data$year <- NA

  # Initialize year and counter
  year <- start_year
  counter <- 0

  # Loop over rows
  for(i in seq_len(nrow(wnba_data))) {
    # If the next rank is 1 and is not NA, increment the counter
    if(i < nrow(wnba_data) && !is.na(wnba_data$Rank[i+1]) && wnba_data$Rank[i+1] == 1) {
      counter <- counter + 1
    }
    # Calculate the year based on the counter
    wnba_data$year[i] <- start_year + counter
  }

  # Remove rows with "Totals" or "Averages"
  wnba_data <- wnba_data[!rowSums(sapply(wnba_data, function(x) grepl("Totals|Averages", x))), ]

  wnba_data <- wnba_data %>%
    mutate(Team = str_sub(Team, start = 1, end = 3)) %>%
    mutate(Team = case_when(
      Team == "NY\n" ~ "NYL",
      Team == "LA\n" ~ "LAX",
      Team == "LV\n" ~ "LVA",
      TRUE ~ Team
    ))

  return(wnba_data)
}

scrape_wnbadata(2024)

# Scrape data for all years
wnba <- bind_rows(scrape_wnbadata(2018),
          scrape_wnbadata(2019),
          scrape_wnbadata(2020),
          scrape_wnbadata(2021),
          scrape_wnbadata(2022),
          scrape_wnbadata(2023),
          scrape_wnbadata(2024))

# Rename columns
colnames(wnba)[1] <- "rank"
colnames(wnba)[2] <- "team"
colnames(wnba)[3] <- "players_active"
colnames(wnba)[4] <- "avg_age_team"
colnames(wnba)[5] <- "total_cap_all"
colnames(wnba)[6] <- "cap_space_all"
colnames(wnba)[7] <- "active"

# Select columns and add new column
wnba <- wnba |>
  select(year, rank, team, players_active, avg_age_team, total_cap_all, cap_space_all, active) |>
  mutate(WM = "W")

wnba
```

```{r, include=FALSE, message=FALSE}
# Create a data frame with team abbreviations, full names, and cities
wnba_teams <- tibble(
  team = c("ATL", "CHI", "CON", "IND", "NYL", "WAS", "DAL", "LVA", "LAX", "MIN", "PHX", "SEA"),
  team_name = c("Atlanta Dream", "Chicago Sky", "Connecticut Sun", "Indiana Fever", "New York Liberty", "Washington Mystics", "Dallas Wings", "Las Vegas Aces", "Los Angeles Sparks", "Minnesota Lynx", "Phoenix Mercury", "Seattle Storm"),
  city = c("Atlanta", "Chicago", "Uncasville", "Indianapolis", "New York", "Washington", "Dallas", "Las Vegas", "Los Angeles", "Minneapolis", "Phoenix", "Seattle")
)

# Join WNBA data with team names and cities
wnba <- wnba %>%
  left_join(wnba_teams, by = "team")
```

```{r, include=FALSE, message=FALSE}
#Combined data

# For NBA data
nba$players_active <- as.character(nba$players_active)

nba$year <- nba$year + 1

# For WNBA data
wnba$players_active <- as.character(wnba$players_active)

# Now bind the rows
all_basketball <- bind_rows(wnba, nba)
```

```{r}
# Generating Map
# Define UI for map tab
latlon <- tabPanel(
  title = "Map",
  leafletOutput("map")
)

# Define server logic for map tab
map_server <- function(input, output) {
  output$map <- renderLeaflet({
    stadiums <- data.frame(
      team = c("Atlanta Hawks", "Boston Celtics", "Brooklyn Nets", "Charlotte Hornets", "Chicago Bulls", "Cleveland Cavaliers", "Dallas Mavericks", "Denver Nuggets", "Detroit Pistons", "Golden State Warriors", "Houston Rockets", "Indiana Pacers", "Los Angeles Clippers", "Los Angeles Lakers", "Memphis Grizzlies", "Miami Heat", "Milwaukee Bucks", "Minnesota Timberwolves", "New Orleans Pelicans", "New York Knicks", "Oklahoma City Thunder", "Orlando Magic", "Philadelphia 76ers", "Phoenix Suns", "Portland Trail Blazers", "Sacramento Kings", "San Antonio Spurs", "Toronto Raptors", "Utah Jazz", "Washington Wizards", "Atlanta Dream", "Chicago Sky", "Connecticut Sun", "Dallas Wings", "Indiana Fever", "Las Vegas Aces", "Los Angeles Sparks", "Minnesota Lynx", "New York Liberty", "Phoenix Mercury", "Seattle Storm", "Washington Mystics"),
      lat = c(33.7573, 42.3662, 40.6826, 35.2251, 41.8807, 41.4964, 32.7905, 39.7487, 42.3410, 37.7680, 29.7508, 39.7640, 34.0430, 34.0430, 35.1382, 25.7814, 43.0450, 44.9795, 29.9490, 40.7505, 35.4634, 28.5392, 39.9012, 33.4456, 45.5316, 38.5805, 29.4269, 43.6435, 40.7683, 38.8981, 33.6451, 41.8538, 41.4896, 32.7291, 39.7640, 36.0905, 34.0430, 44.9795, 40.6826, 33.4456, 47.6222, 38.8684),
      lon = c(-84.3963, -71.0621, -73.9754, -80.8392, -87.6742, -81.6882, -96.8103, -105.0077, -83.0551, -122.3877, -95.3621, -86.1555, -118.2673, -118.2673, -90.0505, -80.1870, -87.9170, -93.2760, -90.0812, -73.9934, -97.5151, -81.3839, -75.1720, -112.0712, -122.6668, -121.4997, -98.4375, -79.3791, -111.9011, -77.0209, -84.4585, -87.6195, -72.0863, -97.1125, -86.1555, -115.1750, -118.2673, -93.2760, -73.9754, -112.0712, -122.3545, -76.9930)
    )

    # Group teams that share the same stadium together
    stadiums <- stadiums %>%
      group_by(lat, lon) %>%
      summarise(team = paste(team, collapse = "<br>"))

    # Create a basketball icon
    basketball_icon <- makeIcon(
      iconUrl = "https://emojiguide.org/images/emoji/6/jvgrthwxjlu6.png",
      iconWidth = 28,
      iconHeight = 28
    )

    # Create a leaflet map
    leaflet(stadiums) %>%
      addTiles() %>%
      addMarkers(~lon, ~lat, icon = basketball_icon, popup = ~team)
  })
}
```

```{r}
# Define UI for data tab
data_ui <- tabPanel(
  title = "Team Statistics",
  sidebarLayout(
    sidebarPanel(
      pickerInput("team_nba", "Select NBA team:", choices = NULL, options = list(`actions-box` = TRUE), multiple = TRUE),
      pickerInput("team_wnba", "Select WNBA team:", choices = NULL, options = list(`actions-box` = TRUE), multiple = TRUE),
      pickerInput("year", "Select year:", choices = NULL, options = list(`actions-box` = TRUE), multiple = TRUE),
      pickerInput("column1", "Select first column:", choices = c("players_active", "avg_age_team", "total_cap_all", "cap_space_all", "active"), options = list(`actions-box` = TRUE)),
      pickerInput("column2", "Select second column:", choices = c("players_active", "avg_age_team", "total_cap_all", "cap_space_all", "active"), options = list(`actions-box` = TRUE))
    ),
    mainPanel(
      fluidRow(
        column(6, h3("NBA"), DT::dataTableOutput("table_nba")),
        column(6, h3("WNBA"), DT::dataTableOutput("table_wnba"))
      )
    )
  )
)

# Define server logic for data tab
data_server <- function(input, output, session) {
  # Update team and year choices
  updatePickerInput(session, "team_nba", choices = unique(all_basketball$team_name[all_basketball$WM == 'M']))
  updatePickerInput(session, "team_wnba", choices = unique(all_basketball$team_name[all_basketball$WM == 'W']))
  updatePickerInput(session, "year", choices = unique(all_basketball$year))

  output$table_nba <- DT::renderDataTable({
    nba_data <- nba
    if (!is.null(input$year)) {
      nba_data <- nba_data[nba_data$year %in% input$year, ]
    }
    if (!is.null(input$team_nba)) {
      nba_data <- nba_data[nba_data$team_name %in% input$team_nba, ]
    }
    nba_data <- nba_data[, c("year", "team_name", input$column1, input$column2)]
    DT::datatable(nba_data)
  })

  output$table_wnba <- DT::renderDataTable({
    wnba_data <- wnba
    if (!is.null(input$year)) {
      wnba_data <- wnba_data[wnba_data$year %in% input$year, ]
    }
    if (!is.null(input$team_wnba)) {
      wnba_data <- wnba_data[wnba_data$team_name %in% input$team_wnba, ]
    }
    wnba_data <- wnba_data[, c("year", "team_name", input$column1, input$column2)]
    DT::datatable(wnba_data)
  })
}
```

```{r}
# Define UI
ui <- navbarPage(
  title = "Location WNBA and NBA Teams",
  latlon,
  data_ui
)

# Define server logic
server <- function(input, output, session) {
  map_server(input, output)
  data_server(input, output, session)
}

# Run the application 
shinyApp(ui = ui, server = server)
```

Conclusion
==========================
Our final project, titled "WNBA and NBA Final," provided an in-depth analysis of professional basketball statistics, focusing on comparing the WNBA and NBA. Utilizing data scraped from various sources such as ESPN, HoopsHype, and Spotrac, we gathered comprehensive datasets for both leagues spanning multiple years. This allowed us to explore various statistical metrics, including player demographics, team performance, and financial expenditures.

#### Key Insights and Findings
Player Demographics:

+ We examined the demographics of WNBA players, focusing on age, height, and position. Our analysis revealed that younger players often occupy forward positions, while no significant relationships were found between age, height, and other statistics.
The interactive dashboard enabled users to filter data by year and various statistical metrics, providing a clear visual representation of the relationships between these demographics and player performance.
Team Performance:

+ Our analysis of team statistics over time showed fluctuations in team performance, particularly in the number of wins and losses. Notably, the 2020 season saw a significant drop in the number of games due to external factors, impacting team performance metrics.
The interactive plots allowed users to explore how different teams performed each year, highlighting trends and changes in performance over time.

Comparison Between WNBA and NBA:
+ By comparing the WNBA and NBA data, we uncovered key differences and similarities between the leagues. For instance, the NBA generally has higher financial expenditures and a larger number of active players due to its longer season and greater commercial reach.

+ Our analysis highlighted the differences in playing style and strategic focus between the two leagues, emphasizing the WNBA's emphasis on teamwork and fundamentals compared to the NBA's fast-paced, star-driven gameplay.

Geographical Distribution:
+ Using Leaflet maps, we visualized the geographical distribution of WNBA and NBA teams. This provided a spatial perspective on where teams are located across the United States, offering insights into regional representation in professional basketball.

#### Future Directions
This project serves as a foundation for further exploration into the dynamics of professional basketball. Potential future directions could include:
+Analyzing the impact of external factors such as injuries, trades, and draft picks on team performance and player statistics.
+Expanding the scope to include international leagues and their comparison with the WNBA and NBA.
+Incorporating advanced statistical models and machine learning techniques to predict player and team performance based on historical data.

#### Conclusion
Through our comprehensive analysis and interactive visualizations, we have provided valuable insights into the world of professional basketball. Our project not only highlights the statistical differences and similarities between the WNBA and NBA but also underscores the importance of data-driven analysis in understanding and appreciating the complexities of these sports. We hope our work inspires further research and exploration into the fascinating world of basketball statistics.

The disparity in the availability of historical statistics for women's basketball, which only began complete record-keeping in 2018 compared to much earlier for men, underscores the need for greater focus and resources in women's sports. With rising fan engagement following events like March Madness and the contributions of standout players like Caitlin Clark and the Aces' A'ja Wilson, we anticipate increased ticket sales and economic stimulation.

Thank you for engaging with our project, and we hope you found it both informative and enjoyable.